---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# gargoyle

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

[WIP]

The goal of gargoyle is to provide a non-reactive, event-based framework for Shiny.

## Installation

You can install the released version of gargoyle from [Github](https://CRAN.R-project.org) with:

``` r
remotes::install_github("ColinFay/gargoyle")
```

## About

`{gargoyle}` is a package that provide wrappers around some Shiny elements to turn your app into and event-based application instead of a full reactive app. The framework is centered around `listen` / `trigger` and `on`.

It works with classical UI, and just needs tweaking the server side of your app.

### What the heck?

Shiny's default reactive behavior is very helpful when it comes to building small applications. Because, you know, the good thing about reactivity is that when something moves somewhere, it's updated everywhere. But the bad thing about reactivity is that when something moves somewhere, it's updated everywhere. So it does work pretty well on small apps, but can get very complicated on bigger apps, and can quickly get out of hands. 

That's where `{gargoyle}` comes: it provides an event based paradigm for building your apps, so that things happen under a control flow. 

### For whom?

If you're just building small shiny apps, you're probably good with Shiny default reactive behavior. But if ever you've struggled with reactivity on more bigger apps, you might find `{gargoyle}` useful. 

### The trade-off

`{gargoyle}` will be more verbose and will demand more work upfront to make things happen. I believe this is for the best if you're working on a big project. 

## Design pattern

A `{gargoyle}` has: 

+ `init`, `listen` & `trigger`, which allow to initiate, listen on, and trigger an event

+ `on`, that runs the `expr` when the event in triggered

## Example

```{r eval = FALSE}
library(shiny)
pkgload::load_all()
ui <- function(request){
  tagList(
    h4('Go'),
    actionButton("y", "y"),
    h4('Output of z$v'),
    tableOutput("evt")
  )
}

server <- function(input, output, session){
  
  init( "plop", "pouet", "poum")
  
  z <- new.env()
  
  observeEvent( input$y , {
    trigger("plop")
    z$v <- mtcars
  })
  
  on("plop", {
    z$v <- airquality
    trigger("pouet")
  })
  
  on("pouet", {
    z$v <- iris
    trigger("poum")
  })
  
  output$evt <- renderTable({
    watch("poum")
    head(z$v) 
  })
  
}

shinyApp(ui, server)
```

You can then get / clear the logs of the times the triggers were called: 

```{r}
get_gargoyle_logs()
clear_gargoyle_logs()
```


<br>

Please note that the 'gargoyle' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
